---
title: 'Homework #1: SDGB-7844'
author: "Chris Hayduk"
fordham_id_number: "A11146981"
fordham_email: "chayduk@fordham.edu"
date: "10/9/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Each of the exercises below are from our textbook - _Introduction to Scientific Programming and Simulaiton Using R._ The exercises begin on p. 26.  Make sure to comment your code.  Points will be deducted for uncommented code!  

You should create new code blocks if necessary to answer the question.  You should also use plaintext to provide an explanation for your code if the question calls for one.

### Chapter 2 - Exercise 1a. 

```{r}


```

### Chapter 2 - Exercise 1b. 

```{r}


```

### Chapter 2 - Exercise 1c. 

```{r}
 

```

### Chapter 2 - Exercise 1d. 

```{r}


```

### Chapter 2 - Exercise 1e. 

```{r}


```


### Chapter 2 - Exercise 2a.

```{r}


```

### Chapter 2 - Exercise 2b.

```{r}


```


### Chapter 2 - Exercise 2c. 

```{r}


```

### Chapter 2 - Exercise 2d.

```{r}
 

```

### Chapter 2 - Exercise 4

```{r}

```


### Chapter 2 - Exercise 5a. 

```{r}

```

### Chapter 2 - Exercise 5b. 

```{r}

```

### Chapter 2 - Exercise 5c.

```{r}

```

### Chapter 2 - Exercise 5d. 

```{r}

```

### Chapter 2 - Exercise 5e. 

```{r}

```

### Chapter 2 - Exercise 6

```{r}

```

### Chapter 2 - Exercise 7

```{r}

```
